---
title: "Tidy TuesdayWeek 14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggthemes)
library(knitr)
library(kableExtra)
library(maps)
#library(gganimate)
library(animation)
library(scales)
```

Tidy Tuesday is a data project hosted every week by the [R for Data Science learning community.](https://github.com/rfordatascience/tidytuesday) Anyone is welcome to participate, and are encouraged to colloborate and share their findings on Twitter with the [R4DS community](https://twitter.com/R4DScommunity).This week we are examining global life expectancy from [this article](https://ourworldindata.org/life-expectancy).  


```{r, echo = FALSE, include=FALSE}
life_exp <- read_csv("/Users/Joe/Documents/Github/tidytuesday/data/week14.csv")
```

## Data Cleaning
In order to use the maps library built in mapping data, the country names had to
match one another in order to be joined correctly. 

```{r}
life_exp$country <- ifelse(life_exp$country == "United States", "USA", life_exp$country)
life_exp$country <- ifelse(life_exp$country == "Democratic Republic of Congo", "Democratic Republic of the Congo", life_exp$country)
life_exp$country <- ifelse(life_exp$country == "Cote d'Ivoire", "Ivory Coast", life_exp$country)
life_exp$country <- ifelse(life_exp$country == "Congo", "Republic of Congo", life_exp$country)
life_exp$country <- ifelse(life_exp$country == "United States", "USA", life_exp$country)
life_exp$country <- ifelse(life_exp$country == "United Kingdom", "UK", life_exp$country)
life_exp$country <- ifelse(life_exp$country == "Serbia (including Kosovo)", "Serbia", life_exp$country)
```

# Making a Table to Highlight Global Percent Change

I thought it would be interesting to examine which countries saw the biggest and
the smallest improvements in life expectancy. I first separated the data into two 
new data frames, one for the year 1950 and opne for the year 2015. I then joined 
the two together and found the difference for every country's change in life 
expectancy and then also calculated the percent change.

```{r}
life_1950 <- life_exp %>% 
  group_by(country) %>% 
  filter(year == 1950)

life_2015 <- life_exp %>% 
  group_by(country) %>% 
  filter(year == 2015)

exp_change <- life_1950 %>% 
  bind_cols(life_2015) %>% 
  mutate(change = round(life_expectancy1 - life_expectancy, 2)) %>% 
  mutate(percent_change = round(change / life_expectancy * 100, 2)) %>% 
  arrange(desc(percent_change))

ten_highest <- exp_change %>% 
  ungroup() %>% 
  select(country, change, percent_change) %>% 
  top_n(10) 

ten_lowest <- exp_change %>% 
  ungroup() %>% 
  select(country, change, percent_change) %>% 
  top_n(-10) %>% 
  arrange(percent_change)

top_10 <- bind_cols(ten_highest, ten_lowest)

top_10$change <- paste("+", top_10$change)
top_10$change1 <- paste("+", top_10$change1)
top_10$percent_change <- paste0(top_10$percent_change, "%")
top_10$percent_change1 <- paste0(top_10$percent_change1, "%")

top_10 %>% 
  kable(caption = "Top Ten Countries with the Largest and Smallest Percent Change in Life Expectancy (1950 - 2015)",
        col.names = c("Country", "Difference in Years", "Largest Percent Change",
                      "Country", "Difference in Years", "Smallest Percent Change")) %>% 
  kable_styling() 
```

This table is the final result. To me, this table is super encouraging because no
country had decreases in life expectancy. The countries that changed the most 
are traditionally considered as "developing" countries. This speaks volumes in 
terms of improvements in available health care and overall growth as countries. 
The majority of the countries with the smallest percent change are countries with
solid health care, so it makes sense that they didnt see large increases in this 
period. This table is neat to show the overall picture, but I think it's weakness
lies in the sense that it doesn't show  what happened betweened these years, just 
the two end points. 

# Creating a Choropleth World Map
Using the data from above, I wanted to create a global map showing these increases 
to help visualize the entire picture, not just a select number of countries.




## Creating the map
Nothing fancy here, I just joined the mapping data with our dataset. 

```{r}
map.world <- map_data("world")
map.world <- left_join(map.world, exp_change, by = c("region" = "country"))

map <- ggplot() +
  geom_polygon(data = map.world,
               aes(x = long, 
                   y = lat,
                   group = group, 
                   fill = percent_change),
               color = "white",
               size = 0.25) +
  scale_fill_gradient(low = "light grey", 
                      high = "dark green",
                      na.value = "white",
                      guide = "colorbar") +
  labs(title = "Global Percent Change in Life Expectancy (1950 - 2015)",
       caption = "@Joe_Stoica") +
  theme(plot.title = element_text(hjust = 0.5, size = 15),
        legend.title = element_blank(), 
        legend.position = "top",
        legend.text = element_text(size = 12),
        panel.background = element_blank(),
        plot.background = element_blank(),
        panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank()
        )
map
```

```{r echo=FALSE, include=FALSE}
ggsave(map, filename = "map.png")
```

# Animating Life Expectancy Across the Entire Time Period
I have never worked with gganimate before, so I thought this would be a good time 
to learn how to use it! Note: As of 7/15/18, I can no longer run this code due to 
gganimate being completely redone. Once the final version is published to CRAN,
I will come back and make this run again. Here is a tweet showing the final
gif though!

```{r, warning=FALSE}
p <- ggplot() +
  geom_density(data = life_exp %>% filter(year >= 1950),
               fill = "#5e8c9e",
               aes(life_expectancy,
                   frame = year, 
                   cumulative = FALSE)) +
  theme_fivethirtyeight() +
  labs(subtitle = "Global Changes in Life Expectancy (1950 -2015)",
       caption = "@Joe_Stoica") + 
  theme(axis.title = element_text())+
  xlab("Life Expectancy") +
  ylab("Density") 

#gganimate(p, interval = 0.1)
```




