---
title: 'Tidy Tuesday Week 13: Global Alcohol Consumption'
author: Joe Stoica
date: 6/29/2018
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
library(tidyverse)
library(ggthemes)
library(knitr)
library(kableExtra)
```

Tidy Tuesday is a data project hosted every week by the [R for Data Science learning community.](https://github.com/rfordatascience/tidytuesday) Anyone is welcome to participate, and are encouraged to colloborate and share their findings on Twitter with the [R4DS community](https://twitter.com/R4DScommunity).This week we are examining global alcohol consumption using the dataset from this 
[Fivethirtyeight article](https://fivethirtyeight.com/features/dear-mona-followup-where-do-people-drink-the-most-beer-wine-and-spirits/). 



```{r read in data, echo=FALSE}
# Windows
#consumption <- read_csv("C:/Users/Joe/Documents/Github/tidytuesday/data/week13_alcohol_global.csv")

# Mac
consumption <- read_csv("/Users/Joe/Documents/Github/tidytuesday/data/week13_alcohol_global.csv")
```

# Examining the Data
```{r}
head(consumption)
```

The above table shows the first six rows in our data frame. We are given a country,
and the number of servings of beer, spirits, and wine consumed by that country, 
as well as the total liters of pure alcohol. It is important to note that this 
column is not the sum of servings.

# Remaking the Table from the 538 Article
I wanted to start this week off by recreating the table shown in the article, which
displays the top ten countries that consumes the most beer, spirits, and wine. 

The first step is to separate our data frame into three smaller data frames by 
drink, get the top ten highest serving values, and then sort them from largest 
to smallest.

```{r 538 table remake}
beer <- consumption %>% 
  select(country, Beer = beer_servings) %>% 
  top_n(10) %>% 
  arrange(desc(Beer))

spirit <- consumption %>% 
  select(country, Spirit = spirit_servings) %>% 
  top_n(10) %>% 
  arrange(desc(Spirit))

wine <- consumption %>% 
  select(country, Wine = wine_servings) %>% 
  top_n(10) %>% 
  arrange(desc(Wine))
```

Once we have those, we bind those three data frames into one and create our table 
using the kable function.

```{r}
alcohol <- bind_cols(beer, spirit, wine)

alcohol %>% 
  kable(caption = "Who Drinks The Most Beer, Spirits, And Wine?",
        col.names = c("", "BEER", "", "SPIRIT", "", "WINE")) %>% 
  kable_styling() 
```



# Highest Pure Alcohol Consumption
I thought it would be interesting to look at pure alcohol consumption next. 
I followed a similar process as above, grabbing the top 20 countries with the
highest pure alcohol consumption.

```{r top 20 plot}

top20 <- consumption %>% 
  top_n(20) %>% 
  select(country, Beer = beer_servings, Spirits = spirit_servings, Wine =  wine_servings) %>% 
  gather(key = drink, value = qty, -country) %>% 
  group_by(country)
```

Now that the data is tidied up, I thought a stacked bar chart would be a great 
way to display not only pure alcohol consumption, but also which kind of alcohol
comprises that countries consumption.

```{r}
plot <- ggplot(top20, aes(reorder(country, qty),
                          qty,
                          fill = drink)) +
  geom_bar(stat = "identity")+
  theme_fivethirtyeight() +
  coord_flip() +
  theme(axis.title = element_text(),
        axis.title.y = element_blank(),
        plot.title = element_text(hjust = 1),
        legend.position = "top",
        legend.text = element_text(size = 14)) +
  ylab("Servings") +
  ggtitle("Top 20 Consumers of Pure Alcohol By Drink Type") +
  guides(fill = guide_legend(keywidth = 3,
                             keyheight = .75, 
                             title=NULL, 
                             label.position = "top", 
                             label.hjust = 0.5,
                             reverse = TRUE)) +
  scale_fill_manual(values = c("#0392cf", "#7bc043", "#ee4035"))

plot
```

```{r, echo=FALSE}
ggsave("top20.png", plot)
```

We now have this lovely graphic. It has the highest consumer (Andorra) on top, 
and the lowest (the UK) on bottom, as well as their servings of alcohol. I think 
it is very interesing how each country is unique in how it drinks, because there
are few similarities in the distribution of servings. 

# Sankey Diagram

This is very similar to above, however I looked at the top ten countries instead 
of 20. Sankey Diagrams are great for showing how quantities flow from one location 
to the next. This graphic is unable to be displayed in this 
document, but you can find it [here](rmd/sankey.html).

```{r sankey}
library(devtools)
install_github("mages/googleVis")
require(googleVis)

top10 <- consumption %>% 
  top_n(10) %>% 
  select(country, Beer = beer_servings, Spirits = spirit_servings, Wine =  wine_servings) %>% 
  gather(key = drink, value = qty, -country) %>% 
  group_by(country)

sankey <- plot(gvisSankey(top10,
                          from = "country",
                          to = "drink",
                          weight = "qty",
                          option = "list"(
                            height = 750,
                            sankey = "{link:{color:{fill:'light blue'}}}"
                          )))
```



